---
title: "Dominance Plots"
author: "Andreas M. Brandmaier and Øystein Sørensen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{domination}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

num.datasets <- 5
```

## Simulation

```{r}
library("metagam")
```

Create `r num.datasets` datasets from standard gam example #1. The noise parameter is scaled, such that the first GAM has lowest noise and the last has highest noise. In a domination plot, we plot the relative contributions of each dataset on the prediction. By the way we generated the data, we expect largest influence by dataset 1 and lowest by dataset `r num.datasets`.

```{r cars}
## simulate datasets
set.seed(123)
datasets <- lapply(1:num.datasets, 
                   function(x) mgcv::gamSim(scale = x, verbose = FALSE))
```

Delete all data in the first dataset that has values lower than .2 on the dimension `x2`. Thus, we expect a low (ie., zero) contribution of the first dataset on low values of dimensions `x2`.

```{r}
df <- datasets[[1]]
df[df$x2<0.2,] <- NA
datasets[[1]] <- df
```

Next, delete all values of the second dataset for large values (`x2`>.8).

```{r}
df <- datasets[[2]]
df[df$x2 > 0.8, ] <- NA
datasets[[2]] <- df
```

## Fit GAMs

Next, we fit the indivial GAMs for each of the eight datasets.

```{r}
## fit a generalized additive model to each dataset separately
models <- lapply(datasets, function(dat){
  ## Full fit using mgcv
  gamfit <- mgcv::gam(y ~ s(x0, bs = "cr") + s(x1, bs = "cr") + 
                        s(x2, bs = "cr"), data = dat)
  ## Extract the necessary components for performing a meta-analysis
  ## This removes all subject-specific data
  strip_rawdata(gamfit)
})
```

## Meta-Analysis

Then, we analyze the term `s(x2)`.

```{r}
meta_analysis <- metagam(models, grid_size = 500, 
                         terms = "s(x2)", intercept = TRUE)
```

## Domination Analysis

Finally, we create a domination plot that tells us how much the points on axis `x2` are 
influenced by the individual GAMs. We see that the influence of the GAMs is graded 
according to the simulated noise levels. Second, we see that on the left-hand side, the influence of the
first dataset is almost zero whereas the influence of the second dataset is almost zero
on the righthand side.

```{r}
plot_domination(meta_analysis)
```

