---
title: "Heterogeneity Plots"
author: "Andreas M. Brandmaier and Øystein Sørensen"
date: "`r Sys.Date()`"
bibliography: ../inst/REFERENCES.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Heterogeneity Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# loads all new features
devtools::load_all(".")
```

The metagam package offers a way to visualize the heterogeneity of the estimated smooth functions over the range of explanatory variables. This will be illustrated here.

## Simulation


```{r setup, eval=FALSE}
library(metagam)
```


We start by simulating 5 datasets using the `gamSim()` function from mgcv. We use the response $y$ and the explanatory variable $x_{2}$, but add an additional shift $\beta x_{2}^{2}$ where $\beta_{2}$ differs between datasets, yielding heterogenous data.

```{r}
library(mgcv)
set.seed(1233)
shifts <- c(0, .5, 1, 0, -1)
datasets <- lapply(shifts, function(x) {
  ## Simulate data
  dat <- gamSim(scale = .1, verbose = FALSE)
  ## Add a shift
  dat$y <- dat$y + x * dat$x2^2
  ## Return data
  dat
})
```

## Fit GAMs

Next, we analyze all datasets, and strip individual participant data.

```{r}
models <- lapply(datasets, function(dat){
  b <- gam(y ~ s(x2, bs = "cr"), data = dat)
  strip_rawdata(b)  
})
```

## Meta-Analysis

Next, we meta-analyze the models. Since we only have a single smooth term, we use `type = "response"` to get the response function. This is equivalent to using `type = "iterms"` and `intercept = TRUE`.

```{r}
meta_analysis <- metagam(models, type = "response")
```

Next, we plot the separate estimates together with the meta-analytic fit. We clearly see that dataset 3, which had a positive shift $\beta=1 x_{2}^2$, lies above the others for $x_{2}$ close to 1, and opposite for dataset 5.

```{r}
plot(meta_analysis)
```

The plotting function for the meta-analysis object is a ggplot2-object, and can thus be altered using standard ggplot syntax. To learn more about customization of ggplot2-objects, please [see the ggplot2 documentation](https://ggplot2.tidyverse.org/).

```{r}
plot(meta_analysis) + 
  ggplot2::scale_colour_brewer(palette = "Set1")
```


We can investigate this further using a heterogeneity plot, which visualizes Cochran's Q-test (@Cochran1954) as a function of $x_{2}$. By default, the test statistic (Q), with 95 \% confidence bands, is plotted. We can see that the confidence band for Q is above 0 for $x_{2}$ larger than about 0.7.

```{r}
plot_heterogeneity(meta_analysis)
```

We can also plot the $p$-value of Cochran's Q-test. The dashed line shows the value $0.05$. The $p$-value plot is in full agreement with the Q-statistic plot above: There is evidence that the underlying functions from each dataset are different for values from about 0.7 and above.

```{r}
plot_heterogeneity(meta_analysis, type = "p")
```




# References
